Why Is Neuron Modeling of Particular Philosophical Interest?

### 2023-01-26 REV_Polak1/Polak1.tex
first Polak review added

\begin{document}
\title{Why Is Neuron Modeling of Particular Philosophical Interest?}
\maketitle

Andrzej Bielecki, \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}, Springer International Publishing, Cham 2019, pp. 156; Studies in Computational Intelligence Volume vol. 770, DOI: \href{https://doi.org/10.1007/978-3-319-90140-4}{10.1007/978-3-319-90140-4}.

A~peculiarity of ``philosophy in science''
%\label{ref:RNDSoIfWpdNhG}(see Heller, 2019; Polak, 2019)
\parencites[see][]{heller_how_2019}[][]{polak_philosophy_2019} %
 is that the best sources tend to be atypical from the viewpoint of most philosophers: For example, on the one hand, there are works that popularize science, while on the other hand, there are research articles and even specialized monographs. The book discussed here falls into the last category, and it is devoted to modeling neurons and perceptrons. It was written by a~mathematician from Kraków, Andrzej Bielecki, who is currently working at the AGH University of Science and Technology.\footnote{Andrzej Bielecki received an M.Sc. degree in Physics and Mathematics and a~Ph.D. in Mathematics, D.Sc. (habilitation) in Mathematics from the Jagiellonian University in Kraków. He obtained a~professorship in Computer Science in 2020. His fields of interest includes dynamical systems theory, artificial intelligence, cybernetics, and philosophy of science, and he has written over 120 scientific papers and one textbook.} Readers of \textit{ZFN}, as well as the related \textit{Semina Scientiarum} journal, will probably associate him with the philosophical activities that he has practiced within the context of his scientific activities 
%\label{ref:RNDkz8F1S0BJI}(Bielecki, 2016; 2018).
\parencites*[][]{bielecki_cybernetyczna_2016}[][]{}. %
 Bielecki is an example of a~scientist–philosopher from the Krakow milieu,\footnote{It is worth mentioning that in the book, Bielecki mostly uses examples from research conducted in the Kraków milieu. The use of the cybernetic theory framework could also be interpreted as another sign of the local milieu's influence.} and it is worth noting that he develops his philosophical activities, among other things, through his work on the Committee on Philosophy of Science at the Polish Academy of Arts and Sciences in Kraków.

@@

Now, let me illustrate the conceptual scheme used by Bielecki: It is based on concepts from cybernetics theory, one of the vital mathematical theories that provides the foundation for developing interdisciplinary research and computational modeling. In Poland, cybernetics is still successfully pursued, especially in Kraków at the AGH University of Science and Technology,\footnote{In private correspondence, the author stated that the most important sources of inspiration on the issue of cybernetics are the works of Tadeusiewicz
%\label{ref:RNDr2UOU7mz1W}(1994; 2009),
\parencites**[][]{tadeusiewicz_problemy_1994}[][]{tadeusiewicz_neurocybernetyka_2009}, %
 who is a~distinguished researcher and the founder of a~vivid center of biocybernetics at AGH in Krakow. A~further source of inspiration were the works of another Krakow scientist, Mariusz Flasiński 
%\label{ref:RNDOF7n6kKWVG}(1997; 2016),
\parencites**[][]{flasinski_every_1997}[][]{flasinski_introduction_2016}, %
 who is affiliated with Jagiellonian University in Kraków. } but contemporary international discussions use somewhat different conceptual systems. A~good example is Gordana Dodig-Crnkovic's article in an issue of ZFN 
%\label{ref:RND95aA53ESTH}(Dodig-Crnkovic, 2022).
\parencite[][]{dodig-crnkovic_search_2022}. %
 The deep analogies between the two approaches are surprising. For example, take Bielecki's phrase: ``Each type of biological cells, including the simplest bacteria, receives stimuli from its environment and processes the obtained signals'' 
%\label{ref:RNDfjWEt2XKUt}(Bielecki, 2019, p.5).
\parencite[][p.5]{bielecki_models_2019}. %
 It is close to the info-computational in Dodig-Crnkovic's view, although she uses a~specific reference to information theory. It would be worthwhile to analyze the relationship between cybernetics and contemporary information concepts in more depth, because it may be possible to find new, inspiring analogies or more convenient conceptual frameworks.

@@

\section{Bibliography}
Bielecki, A., 2016. Cybernetyczna analiza zjawiska życia. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (61), pp.133–164. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/361{\textgreater} [Accessed 27 January 2020].

Bielecki, A., 2018. Epistemologiczne problemy w~biologii subkomórkowej: obserwacje, modele matematyczne i~symulacje komputerowe. \textit{Semina Scientiarum}, [online] 16, pp.10–23. https://doi.org/10.15633/ss.2482.

Bielecki, A., 2019. \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}. Studies in Computational Intelligence. [online] Cham: Springer International Publishing. https://doi.org/10.1007/978-3-319-90140-4.

Dodig-Crnkovic, G., 2022. In search of common, information-processing, agency-based framework for anthropogenic, biogenic, and abiotic cognition and intelligence. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), (73), p.??-??

Flasiński, M., 1997. ‘Every Man in His Notions' or Alchemists' Discussion on Artificial Intelligence. \textit{Foundations of Science}, [online] 2(1), pp.107–121. https://doi.org/10.1023/A:1009687513096.

Flasiński, M., 2016. \textit{Introduction to Artificial Intelligence}. [online] Cham: Springer International Publishing. https://doi.org/10.1007/978-3-319-40022-8.

Heller, M., 2019. How is philosophy in science possible? \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (66), pp.231–249. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/482{\textgreater} [Accessed 6 October 2021].

Kycia, R., 2021. Information and brain. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (70), pp.45–72. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/514{\textgreater} [Accessed 6 December 2022].

Polak, P., 2019. Philosophy in science: A~name with a~long intellectual tradition. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (66), pp.251–270. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/472{\textgreater} [Accessed 6 October 2021].

Tadeusiewicz, R., 1994. \textit{Problemy biocybernetyki}. Wyd. 2 ed. Warszawa: Wydawnictwo Naukowe PWN.

Tadeusiewicz, R. ed., 2009. \textit{Neurocybernetyka teoretyczna}. Warszawa: Wydawnictwa Uniwersytetu Warszawskiego.

Weisberg, J., 2010. Bootstrapping in General. \textit{Philosophy and Phenomenological Research}, [online] 81(3), pp.525–548. Available at: {\textless}https://www.jstor.org/stable/41057492{\textgreater} [Accessed 6 December 2022].

\section{Abstract}
This review article discusses Andrzej Bielecki's book \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}, as published by Springer International Publishing. This work exemplifies ``philosophy in science'' by adopting a~broad, multidisciplinary perspective for the issues related to the simulation of neurons and neural networks, and the author has addressed many of the important philosophical assumptions that are entangled in this area of modeling. Bielecki also raises several important methodological issues about modeling. This book is recommended for any philosophers who wish to learn more about the current state of neural modeling and find inspiration for a~deeper philosophical reflection on the subject.

\section{Keywords}
neuron modeling, sub-neuron modeling, computational modeling, analog computation, philosophy in science, philosophy of biology, philosophy of computing

\end{document}

>>>>

\begin{newrevengenv}{Paweł Polak}
	{Why is neuron modeling of particular philosophical interest?}
	{Why is neuron modeling of particular philosophical interest?}
	{Why is neuron modeling of particular philosophical interest?}
	{Center for Theoretical Physics, Polish Academy of Sciences}
	{Andrzej Bielecki, \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}, Springer International Publishing, Cham 2019, pp. 156; Studies in Computational Intelligence Volume vol. 770, DOI: \href{https://doi.org/10.1007/978-3-319-90140-4}{10.1007/978-3-319-90140-4}.}

%\begin{document}
%\title{Why is neuron modeling of particular philosophical interest?}
%\maketitle

A~peculiarity of ``philosophy in science''
%\label{ref:RNDSoIfWpdNhG}(see Heller, 2019; Polak, 2019)
\parencites[see][]{heller_how_2019}[][]{polak_philosophy_2019} %
 is that the best sources tend to be atypical from the viewpoint of most philosophers: For example, on the one hand, there are works that popularize science, while on the other hand, there are research articles and even specialized monographs. The book discussed here falls into the last category, and it is devoted to modeling neurons and perceptrons. It was written by a~mathematician from Kraków, Andrzej Bielecki, who is currently working at the AGH University of Science and Technology.\footnote{Andrzej Bielecki received an M.Sc. degree in Physics and Mathematics and a~Ph.D. in Mathematics, D.Sc. (habilitation) in Mathematics from the Jagiellonian University in Kraków. He obtained a~professorship in Computer Science in 2020. His fields of interest includes dynamical systems theory, artificial intelligence, cybernetics, and philosophy of science, and he has written over 120 scientific papers and one textbook.} Readers of \textit{ZFN}, as well as the related \textit{Semina Scientiarum} journal, will probably associate him with the philosophical activities that he has practiced within the context of his scientific activities 
%\label{ref:RNDkz8F1S0BJI}(Bielecki, 2016; 2018).
\parencites*[][]{bielecki_cybernetyczna_2016}[][]{bielecki_epistemologiczne_2018}. %
 Bielecki is an example of a~scientist–philosopher from the Krakow milieu,\footnote{It is worth mentioning that in the book, Bielecki mostly uses examples from research conducted in the Kraków milieu. The use of the cybernetic theory framework could also be interpreted as another sign of the local milieu's influence.} and it is worth noting that he develops his philosophical activities, among other things, through his work on the Committee on Philosophy of Science at the Polish Academy of Arts and Sciences in Kraków.

@@

Now, let me illustrate the conceptual scheme used by Bielecki: It is based on concepts from cybernetics theory, one of the vital mathematical theories that provides the foundation for developing interdisciplinary research and computational modeling. In Poland, cybernetics is still successfully pursued, especially in Kraków at the AGH University of Science and Technology,\footnote{In private correspondence, the author stated that the most important sources of inspiration on the issue of cybernetics are the works of Tadeusiewicz
%\label{ref:RNDr2UOU7mz1W}(1994; 2009),
\parencites*[][]{tadeusiewicz_problemy_1994}[][]{tadeusiewicz_neurocybernetyka_2009}, %
 who is a~distinguished researcher and the founder of a~vivid center of biocybernetics at AGH in Krakow. A~further source of inspiration were the works of another Krakow scientist, Mariusz Flasiński 
%\label{ref:RNDOF7n6kKWVG}(1997; 2016),
\parencites*[][]{flasinski_every_1997}[][]{flasinski_introduction_2016}, %
 who is affiliated with Jagiellonian University in Kraków. } but contemporary international discussions use somewhat different conceptual systems. A~good example is Gordana Dodig-Crnkovic's article in an issue of ZFN 
%\label{ref:RND95aA53ESTH}(Dodig-Crnkovic, 2022).
\parencite[][]{dodig-crnkovic_search_2022}. %
 The deep analogies between the two approaches are surprising. For example, take Bielecki's phrase: ``Each type of biological cells, including the simplest bacteria, receives stimuli from its environment and processes the obtained signals'' 
%\label{ref:RNDfjWEt2XKUt}(Bielecki, 2019, p.5).
\parencite[][p.5]{bielecki_models_2019}. %
 It is close to the info-computational in Dodig-Crnkovic's view, although she uses a~specific reference to information theory. It would be worthwhile to analyze the relationship between cybernetics and contemporary information concepts in more depth, because it may be possible to find new, inspiring analogies or more convenient conceptual frameworks.

@@

%\section{Bibliography}
%Bielecki, A., 2016. Cybernetyczna analiza zjawiska życia. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (61), pp.133–164. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/361{\textgreater} [Accessed 27 January 2020].
%
%Bielecki, A., 2018. Epistemologiczne problemy w~biologii subkomórkowej: obserwacje, modele matematyczne i~symulacje komputerowe. \textit{Semina Scientiarum}, [online] 16, pp.10–23. https://doi.org/10.15633/ss.2482.
%
%Bielecki, A., 2019. \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}. Studies in Computational Intelligence. [online] Cham: Springer International Publishing. https://doi.org/10.1007/978-3-319-90140-4.
%
%Dodig-Crnkovic, G., 2022. In search of common, information-processing, agency-based framework for anthropogenic, biogenic, and abiotic cognition and intelligence. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), (73), p.??-??
%
%Flasiński, M., 1997. ‘Every Man in His Notions' or Alchemists' Discussion on Artificial Intelligence. \textit{Foundations of Science}, [online] 2(1), pp.107–121. https://doi.org/10.1023/A:1009687513096.
%
%Flasiński, M., 2016. \textit{Introduction to Artificial Intelligence}. [online] Cham: Springer International Publishing. https://doi.org/10.1007/978-3-319-40022-8.
%
%Heller, M., 2019. How is philosophy in science possible? \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (66), pp.231–249. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/482{\textgreater} [Accessed 6 October 2021].
%
%Kycia, R., 2021. Information and brain. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (70), pp.45–72. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/514{\textgreater} [Accessed 6 December 2022].
%
%Polak, P., 2019. Philosophy in science: A~name with a~long intellectual tradition. \textit{Philosophical Problems in Science (Zagadnienia Filozoficzne w~Nauce}), [online] (66), pp.251–270. Available at: {\textless}https://zfn.edu.pl/index.php/zfn/article/view/472{\textgreater} [Accessed 6 October 2021].
%
%Tadeusiewicz, R., 1994. \textit{Problemy biocybernetyki}. Wyd. 2 ed. Warszawa: Wydawnictwo Naukowe PWN.
%
%Tadeusiewicz, R. ed., 2009. \textit{Neurocybernetyka teoretyczna}. Warszawa: Wydawnictwa Uniwersytetu Warszawskiego.
%
%Weisberg, J., 2010. Bootstrapping in General. \textit{Philosophy and Phenomenological Research}, [online] 81(3), pp.525–548. Available at: {\textless}https://www.jstor.org/stable/41057492{\textgreater} [Accessed 6 December 2022].

%\section{Abstract}
%This review article discusses Andrzej Bielecki's book \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}, as published by Springer International Publishing. This work exemplifies ``philosophy in science'' by adopting a~broad, multidisciplinary perspective for the issues related to the simulation of neurons and neural networks, and the author has addressed many of the important philosophical assumptions that are entangled in this area of modeling. Bielecki also raises several important methodological issues about modeling. This book is recommended for any philosophers who wish to learn more about the current state of neural modeling and find inspiration for a~deeper philosophical reflection on the subject.
%
%\section{Keywords}
%neuron modeling, sub-neuron modeling, computational modeling, analog computation, philosophy in science, philosophy of biology, philosophy of computing

\vspace{15mm}%
{\subsubsectit{\hfill Abstract}}\\
{This review article discusses Andrzej Bielecki's book \textit{Models of Neurons and Perceptrons: Selected Problems and Challenges}, as published by Springer International Publishing. This work exemplifies ``philosophy in science'' by adopting a~broad, multidisciplinary perspective for the issues related to the simulation of neurons and neural networks, and the author has addressed many of the important philosophical assumptions that are entangled in this area of modeling. Bielecki also raises several important methodological issues about modeling. This book is recommended for any philosophers who wish to learn more about the current state of neural modeling and find inspiration for a~deeper philosophical reflection on the subject.}\par%
\vspace{2mm}%
{\subsubsectit{\hfill Keywords}}\\
{neuron modeling, sub-neuron modeling, computational modeling, analog computation, philosophy in science, philosophy of biology, philosophy of computing.}%

\end{newrevengenv}
